---
title: "Tres Rios Water Quality Data Analysis"
output: html_notebook
---

This script is used to process Tres Rios water quality data. This includes calculating averages and standard errors for samples at all sampling locations, as well as joining these sampling locations into summary tabels for marsh sampling gradients and whole-system sampling gradients.

```{r HOUSKEEPING}

# libraries ----
library('plyr') # always load plyr before dplyr
library("dplyr")
library("tidyr")
library("stringr")
library("lubridate")

#import and clean agpp data ----
water_quality <- read.csv('/Users/chrissanchez/Documents/weel/tres rios/datasets/tr-water-quality/data/tr_waterquality_fielddata.csv', stringsAsFactors = FALSE)
water_quality[water_quality == ''] <- NA # change all missing values to NA

#convert various date and number formats
water_quality <- water_quality %>%
  mutate(date = mdy(date)) #converts to date format using lubridate

#define a function to compute standard error - note this assumes sample size always 3, need to confirm whether I can assume that
se = function(x)
  (sd(x))/sqrt(3)

```

**Summary Data**
This next chunk will average values for all variables across all samples at a given location (i.e. inflow, outflow) for a given month. 

```{r SUMMARY DATA}

#create a data.table to compute and store summary data from shore samples across all transects
all_data_summary <- water_quality %>%
  mutate(month = month(date)) %>%
  mutate(year = year(date)) %>%
  group_by(month, year, location) %>%
  summarize(
        tn = mean(tn_mgL),
        tn_se = se(tn_mgL),
        tp = mean(tp_mgL),
        tp_se = se(tp_mgL),
        doc = mean(doc_mgL),
        doc_se = se(doc_mgL),
        no2 = mean(no2_mgL),
        no2_se = se(no2_mgL),
        no3 = mean(no3_mgL),
        no3_se = se(no3_mgL),
        nh4 = mean(nh4_mgL),
        nh4_se = se(nh4_mgL),
        po4 = mean(po4_ugL),
        po4_se = se(po4_ugL),
        cl = mean(cl_mgL),
        cl_se = se(cl_mgL),
        cond = mean(conductance_us),
        cond_se = se(conductance_us),
        spec_cond = mean(spec_cond_us),
        spec_cond_se = se(spec_cond_us),
        temperature_c = mean(temperature_c),
        temperature_se = se(temperature_c),
        ph = mean(ph),
        ph = se(ph)
  ) %>%
  #mutate(ph = as.numeric(ph)) %>%
  #mutate(temperature_se = as.numeric(temperature_se)) %>%
  arrange(year, month) %>%
  ungroup()

```

**Location-specific summary data**
This chunk will take the previous chunk's output and split it out into 4 location-specific data tables. In the previous chunk's output ("all_data_summary"), location is stored as variable which will potentially make graphing multiple locations on one plot difficult downstream.

```{r LOCATION-SPECIFIC DATA TABLES}

#marsh shore summary data
marsh_shore_summary <- all_data_summary %>%
  filter(location == "shore") %>%
  select(-(tn:tp_se)) %>% #removes tn/tp columns (not measured for transects)
  select(-(location)) 
colnames(marsh_shore_summary) <- paste("marsh_shore", colnames(marsh_shore_summary), sep = "_") #pastes a prefix to all columns
#remove prefix from first three columns to allow proper data.table joining downstream
names(marsh_shore_summary)[1] <- "month"
names(marsh_shore_summary)[2] <- "year"

#marsh water summary data
marsh_water_summary <- all_data_summary %>%
  filter(location == "water") %>%
  select(-(tn:tp_se)) %>% #removes tn/tp columns (not measured for transects)
  select(-(location)) #removes location (will be specified anyways in variable names)
colnames(marsh_water_summary) <- paste("marsh_water", colnames(marsh_water_summary), sep = "_") #pastes a prefix to all columns
#remove prefix from first two columns to allow proper data.table joining downstream
names(marsh_water_summary)[1] <- "month"
names(marsh_water_summary)[2] <- "year"

#whole-system inflow summary data
inflow_summary <- all_data_summary %>%
  filter(location == "inflow") %>%
  select(-(location)) #removes location (will be specified anyways in variable names)
colnames(inflow_summary) <- paste("inflow", colnames(inflow_summary), sep = "_") #pastes a prefix to all columns
#remove prefix from first two columns to allow proper data.table joining downstream
names(inflow_summary)[1] <- "month"
names(inflow_summary)[2] <- "year"

#whole-system outflow summary data
outflow_summary <- all_data_summary %>%
  filter(location == "outflow") %>%
  select(-(location)) #removes location (will be specified anyways in variable names)
colnames(outflow_summary) <- paste("outflow", colnames(outflow_summary), sep = "_") #pastes a prefix to all columns
#remove prefix from first two columns to allow proper data.table joining downstream
names(outflow_summary)[1:2] <- c("month", "year")

```

**Summary Data Tables**

This chunk joins the two transect-level and two whole-system summary datasets generated previously into tables for transect (shore, water) and whole-system (inflow, outflow) summaries, respectively.

```{r SUMMARY DATA}

transect_summary <- merge(marsh_shore_summary, marsh_water_summary, by = c("month", "year")) %>%
  arrange(year, month) %>%
  #need to create a date column for graphing purposes
  mutate(date = paste("01", month, year, sep = "/")) %>%
  mutate(date = dmy(date)) %>%
  #reorder columns to put date column first
  select(date, everything())

whole_system_summary <- merge(inflow_summary, outflow_summary, by = c("month", "year")) %>%
  arrange(year, month) %>%
  #need to create a date column for graphing purposes
  mutate(date = paste("01", month, year, sep = "/")) %>%
  mutate(date = dmy(date)) %>%
  #reorder columns to put date column first
  select(date, everything())

```

```{r FLUX CALCULATIONS}
#note - need to have run all of the tranpsiration analysis code first to update 

tr_water_budget <- read.csv('/Users/chrissanchez/Documents/weel/tres rios/datasets/tr-water-budget/results/tr_water_budget.csv', stringsAsFactors = FALSE)

#fixing the data up 
tr_nutrient_fluxes <- merge(tr_water_budget, whole_system_summary, by = c("month", "year")) %>%
  arrange(year, month) %>%
  mutate(date = paste("01", month, year, sep = "/")) %>%
  mutate(date = dmy(date)) %>%
  #reorder columns to put date column first
  select(date, everything()) %>%
  #convert waterflux in m3 to L
  mutate(
    monthly_frw1_inflow_L = monthly_frw1_inflow_m3*1000,
    monthly_frw1_outflow_L = monthly_frw1_outflow_m3*1000) %>%
  #get rid of all columns except water flux and nutrients
  select(date, monthly_frw1_inflow_L, monthly_frw1_outflow_L, inflow_tn, inflow_tn_se, inflow_tp, inflow_tp_se, outflow_tn, outflow_tn_se, outflow_tp, outflow_tp_se, inflow_nh4, inflow_nh4_se, outflow_nh4, outflow_nh4_se)
  
  
#calculate fluxes!
tr_nutrient_fluxes <- tr_nutrient_fluxes %>%  
  #tn
  mutate(
    inflow_tn_flux_kg = ((inflow_tn*monthly_frw1_inflow_L)/1000000),
    inflow_tn_flux_se_kg = (inflow_tn_se*monthly_frw1_inflow_L)/1000000,
    outflow_tn_flux_kg = (outflow_tn*monthly_frw1_outflow_L)/1000000,
    outflow_tn_flux_se_kg = (outflow_tn_se*monthly_frw1_outflow_L)/1000000) %>%
  #tp
  mutate(
    inflow_tp_flux_kg = (inflow_tp*monthly_frw1_inflow_L)/1000000,
    inflow_tp_flux_se_kg = (inflow_tp_se*monthly_frw1_inflow_L)/1000000,
    outflow_tp_flux_kg = (outflow_tp*monthly_frw1_outflow_L)/1000000,
    outflow_tp_flux_se_kg = (outflow_tp_se*monthly_frw1_outflow_L)/1000000) %>%
  #nh4
  mutate(
    inflow_nh4_flux_kg = (inflow_nh4*monthly_frw1_inflow_L)/1000000,
    inflow_nh4_flux_se_kg = (inflow_nh4_se*monthly_frw1_inflow_L)/1000000,
    outflow_nh4_flux_kg = (outflow_nh4*monthly_frw1_outflow_L)/1000000,
    outflow_nh4_flux_se_kg = (outflow_nh4_se*monthly_frw1_outflow_L)/1000000)

#use select to clean up spreadsheet
tr_nutrient_fluxes <- tr_nutrient_fluxes %>%
  select(-(monthly_frw1_inflow_L:outflow_nh4_se))





```



**EXPORT RESULTS**

```{r RESULTS EXPORT}

#write certain datasheets as csv to the current working directory

write.csv(whole_system_summary, file = "/Volumes/GoogleDrive/My Drive/weel/tres rios/datasets/tr-water-quality/results/tr_waterquality_whole_system_summary.csv")

write.csv(transect_summary, file = "/Volumes/GoogleDrive/My Drive/weel/tres rios/datasets/tr-water-quality/results/tr_waterquality_transect_summary.csv")

```
 
To do:
1) NEED TO FIX CLASS CHANGES FOR PH/TEMP - not high priority since we don't graph these data

Notes:
20170725
1) Another (potentially easier) way to do this might have been to just generate one big table that summarizes the water quality variables (no3, etc.) and still has transect/location columns. Then do a few chunks using filter() to generate the individual tables (marsh_shore_summary etc.). Then merge them back together again. 
